---
title: "Pakistan Economy at a Glance"
author: "Zahid Asghar"
format: 
  dashboard

---

```{r setup, include=FALSE}
# Load required libraries
library(flexdashboard)
library(ggplot2)
library(readxl)
library(tidyverse)
library(plotly)
library(ggthemes)
if(!require(pacman)) install.packages("pacman")

pacman::p_load(tidyverse,reactable,
  gapminder, plotly,shiny, bsicons,bslib, rnaturalearth, countrycode, htmltools)
```

# HOME

## head {height=70%}

### GDP Trends

```{r}
# Load GDP data
gdp <- read_excel("D:/RepTemplates/Dashboard/gdp.xlsx")

# Create GDP plot
g <- ggplot(gdp, aes(x = Date, y = GDP_constant)) +
  geom_line(color = "blue") +
  labs(
    title = "GDP in Constant Prices (Millions of PKR)",
    caption = "Source: Karandaz"
  ) +
  theme_minimal()

# Render interactive plot
ggplotly(g)
```

## head {height=30%}

### Income Per Capita


```{r}
value_box(
  title = "GDP Per Capita",
  value = 1500,
  showcase = bsicons::bs_icon("currency-dollar"),
  theme = value_box_theme(bg="#518fd6")#,
  #p(paste(highest_life_exp_country))
 )
```


### Inflation Rate

```{r}
value_box(value = 13, showcase = bsicons::bs_icon("cash-stack"),
          theme = value_box_theme(bg = "#f0ad4e")

, title = "Inflation Year on Year Basis Jan-2022", color = "red")
```

### Policy Rate

```{r}
value_box(
  title = "Policy Rate",
  value = "9%",
  showcase = bsicons::bs_icon("percent"),
  theme = value_box_theme(bg = "#f0ad4e") # Orange for moderate concern
)

```

Balance of Trade
================

## head {height=70%}
### Balance of Trade Trends

```{r}
# Load Balance of Trade data
X_M <- read_excel("D:/RepTemplates/Dashboard/Balance of Trade.xlsx")

# Reshape data for visualization
Trade_balance <- X_M %>%
  select(Date, Exports, Imports, BoT) %>%
  pivot_longer(cols = -Date, names_to = "variable", values_to = "value")

# Create trade balance plot
p <- ggplot(Trade_balance, aes(x = Date, y = value, color = variable)) +
  geom_line() +
  scale_color_manual(values = c("darkred", "steelblue", "purple")) +
  labs(
    title = "Pakistan Balance of Trade (Millions of USD)",
    caption = "Source: Karandaz"
  ) +
  theme_minimal()

# Render interactive plot
ggplotly(p)
```

## head {height=30%}

### Trade Deficit FY2021-22

```{r}
value_box(value = 24, showcase=bsicons::bs_icon("currency-dollar") , title = "Trade Deficit in FY21-22 (Billions of USD)", color = "red")
```

### Freelancing Revenue

```{r}
value_box(
  value = 216, 
  showcase = bsicons::bs_icon("globe"), 
  title = "Freelancing Revenue (Millions of USD)", 
  theme = value_box_theme(bg = "#5bc0de")
)
```

### Reflection

Hoping against hope, it's been more than two decades of promises to overcome trade deficits. Improving exports requires significant investment in human capital and governance. Without addressing these challenges, progress in reducing the trade deficit may remain elusive.

Fiscal Report Card: Half-Year22
===============================

Column {data-width=650}
-----------------------------------------------------------------------

### Fiscal Operations

```{r}
# Load fiscal data
fiscal <- read_excel("D:/RepTemplates/Dashboard/Fiscal_operations.xlsx")

# Reshape data for visualization
fiscal_balance <- fiscal %>%
  select(Year, `Total Fiscal Expenditure`, `Net Fiscal Revnue`, `Overall Fiscal Balance`) %>%
  pivot_longer(cols = -Year, names_to = "variable", values_to = "value")

# Create fiscal operations plot
g1 <- ggplot(fiscal_balance, aes(x = Year, y = value, color = variable)) +
  geom_line() +
  scale_color_manual(values = c("darkred", "steelblue", "purple")) +
  labs(
    title = "Half-Year: Federal Fiscal Operations (PKR in Billion)",
    x = "Year",
    y = "PKR (Billion)",
    caption = "Source: Ministry of Finance"
  ) +
  theme_minimal()

# Render interactive plot
ggplotly(g1)
```

### Notes

- Data Source: Ministry of Finance
- Half-year data provides a snapshot of fiscal trends in expenditures and revenues.
```

### Steps to Use
1. Save the file as `pakistan_economy_dashboard.qmd`.
2. Open your R environment and set your working directory to the location of the `.qmd` file.
3. Run the following command in the R console to render the dashboard:
   ```r
   quarto::quarto_render("pakistan_economy_dashboard.qmd")
   ```
4. The rendered dashboard will open in your web browser.

This layout integrates flexdashboard features with tidyverse data manipulation and ggplot2 for visualization, making it interactive and visually appealing.