---
title: "[Karachi-R]{.darkblue}"
title-block-banner: docs/images/R-user.webp
author: "Zahid Asghar"
date: today

format: 
  revealjs:
    toc: true
    code-fold: true
    code-tools: true
    code-link: true
    theme: docs/custom.scss
    chalkboard: true
    
editor: visual
---

## [Quarto]{.yellow}

Quarto enables you to ***weave together content*** and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.

## Typing

Upload tidyverse packages

```{r}
library(tidyverse)
```  

::: {.callout-caution}

This is important to understand why we work with tidyverse packages
:::


## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).



