---
title: "Role of middle class in Pakistan's politics <br> and economy: Is it <br> playing an effective role"
subtitle: "PAS"
logo: "images/soe_logo.jpg"
author: "...."
date: today
format:
  pptx: 
    slide-number: true
    chalkboard: true
    #   buttons: false 
    theme: [simple, custom2.scss]
    title-slide-attributes:
      data-background-color: "#cf3a36"
resources:
  - slides.pdf
---

```{r include=FALSE}
library(tidyverse)
library(httr)
clrs <- MetBrewer::met.brewer(name = "Java")
clrs_lt <- colorspace::lighten(clrs, 0.9)
knitr::opts_chunk$set(fig.retina = 3, collapse = TRUE)
options(digits = 3, width = 75)
```

## What makes middle class important {background-color="`r clrs_lt[5]`"}

Middle class: - Is a precondition of stability in the social structures - Is a measure of mitigating inequality in a society - Is a key factor in forcing politicians and bureaucracy to make accountable

> Easterly(2001): Nations with a large middle class tend to grow faster.

-   This is universal phenomenon.
-   English middle class of 18th and 19th century
-   China and India middle class is a phenomenon of today

##  {background-color="`r clrs_lt[5]`"}

-   Collier (1999) : Middle class forms alliance with lower classes in order to demand an inclusive political system
-   Acemoglu (2003): Decisive voters in democracy are from middle class
-   Bannerjee (2007): middle class is a source of new entrenpreneurs and this class stresses on accumulation of human capital and savings

**Warning:** 
> Basur(2003) : If venues for channeling middle class human capital and savings not provided, then tendency towards consumption expenditure puts sustainability of growth in danger

## How big is middle class in Pakistan? {.center .middle background-color="`r clrs[3]`"}


```{r}
#| echo: false
#| warning: false
library(readxl)
library(gt)
mid_class<-read_excel("middle_class.xlsx",sheet = "Sheet2")
mid_class %>% gt() %>% 
  tab_header(title="Method 1 : Expenditure from PPP $2 to $20 per person per day")
```

## How big is middle class in Pakistan? {.center .middle background-color="`r clrs[3]`"}

```{r}
#| echo: false
#| warning: false
library(readxl)
library(gt)
mid_class1<-read_excel("middle_class.xlsx",sheet = "Sheet1")
mid_class1 %>% gt() %>% 
  tab_header(title="Method 2 : Estimation of Middle Class through a Weighted Composite Index")
```


# The middle class is an urban phenomenon – lives and works in cities {background-color="`r clrs[6]`"}

## What does this middle class want? {background-color="`r clrs_lt[6]`"}

They want:
-   Education
-   Jobs

Nayab(2011):
>	When compared to its neighbours Pakistan has a bigger middle class

>	The middle class in Pakistan has actually grown with time

- Declining growth, Rising UE
- GDP (% growth rate) not sufficient to absorb burgeoning youth

- No significant increase in average years of schooling (age 15-19) and its about 8 years since decades
- Currently no space for middle class (intellectual and social life)

# Accommodating middle class {background-color="`r clrs[5]`"}

## {background-color="`r clrs_lt[5]`"}


-   Urban Growth and Management

-   Births and migration will drive growth

**Labour market shortages:**
-	Design
-	Engineering
-	Middle management


-   Localized savings instruments
-   Opportunity for niche cities
-   Backward linkages with farm activity
-   Diversification towards higher value addition


# Inclusive Governance for Middle Class {background-color="`r clrs[5]`"}
-   Infrastructure
-   governance reform of PSEs
-   Social sector governance 
-   18th Amendment
-   Legal and judicial reforms for inclusive markets
-   Efficiency of public expenditure 
-   results based management




