---
title: "Privatisation"
date: "2024-09-07"
format: 
  typst: 
    slide-number: true
    #toc: true
    theme: simple
    incremental: true
execute:
  echo: false
  warning: false
  message: false
---


# Privatisation in Pakistan

> ## The ultimate goal should be to stop the bleeding of public money.  

# No empirical study to verify or refute the claim that privatisation done in the past was successful and in which sectors it was successful.



### 4TH INDUSTRIAL REVOLUTION (Today)\nCyber Physical Systems & Networks (AI, IoT, Blockchain)

### 3RD INDUSTRIAL REVOLUTION (1969)\nAutomation, Electronics & IT Systems 

### 2ND INDUSTRIAL REVOLUTION (1870)\nMass Production & Electricity

### 1ST INDUSTRIAL REVOLUTION (1784)\nMechanisation & Steam Power


## THREE DOGMAS OF NEOLIBERAL ECONOMICS


### 1. THE STATE HAS NO ROLE IN THE ECONOMY, NOR IN ECONOMIC ENTERPRISE.\n\n

### 2. DISINVEST, PRIVATISE AND DEREGULATE ALL STATE ENTERPRISES.\n\n
### 3. THE 'MARKET' CAN PROVIDE EVERYTHING, LET THE PRIVATE SECTOR DO BUSINESS 'EASILY' (WITH NO OVERSIGHT)."

### Political Decisions for Electricity and Gas Sector Provsision in Pakistan 







25 entities for sale in the first phase of privatisation.

- 207 total entities: 11 contribute 80% of losses.
- 3 are from transport sector, 8 from power sector.

## SOE Profitability Trends 

- **World Bank report**: Declining profitability of SOEs for over a decade.
- Profitability dropped from **0.8% of GDP in 2014** to **losses of 0.4% of GDP in 2020**.
- SOEs in Pakistan now have the lowest profitability in the South Asian region.

## Losses 


```{r}
# Install necessary packages if you haven't already
# install.packages("ggplot2")
# install.packages("treemapify")

# Load the required libraries
library(ggplot2)
library(treemapify)

# Create a data frame with SOEs and their corresponding losses
soes_losses <- data.frame(
  SOE = c("National Highway Authority (NHA)", 
           
          "Quetta Electric Supply Company Limited (QESCO)", 
          "Peshawar Electric Supply Company Limited (PESCO)", 
          "Lahore Electric Supply Company Limited (LESCO)", 
          "Sukkur Electric Power Company Limited (SEPCO)", 
          "Multan Electric Power Company Limited (MEPCO)", 
          "Hyderabad Electric Supply Company Limited (HESCO)", 
          "Faisalabad Electric Supply Company Limited (FESCO)", 
          "Islamabad Electric Supply Company Limited (IESCO)", 
          "Pakistan International Airlines Corporation (PIA)", 
          "Pakistan Railways", 
          "Pakistan Steel Mills Corporation (Private) Limited", 
          "Sui Southern Gas Company Limited (SSGC)", 
          "Pakistan Telecommunication Company Limited (PTCL)", 
          "GENCO-II Central Power Generation Company Limited, Thermal Power Station Guddu", 
          "GENCO-III Northern Power Generation Company Limited, Thermal Power Station Muzaffargarh", 
          "Pakistan Post Office", 
          "Pakistan Broadcasting Corporation (PBC)", 
          "GENCO-I Jamshoro Power Company Limited", 
          "GENCO-IV Lakhra Power Generation Company Limited", 
          "Pakistan Expo Centres (Pvt) Ltd", 
          "Pakistan Revenue Automation (Private) Limited", 
          "State Engineering Corporation (Private) Limited"),
  
  Losses_Rs_Billion = c(413.454, 
                         
                        88.487, 
                        80.595, 
                        30.839, 
                        30.667, 
                        23.372, 
                        15.640, 
                        14.983, 
                        0.666, 
                        75.755, 
                        48.534, 
                        25.455, 
                        17.350, 
                        15.544, 
                        14.170, 
                        3.240, 
                        3.234, 
                        1.617, 
                        0.910, 
                        0.480, 
                        0.181, 
                        0.165, 
                        0.032)
)

# Filter for losses greater than 5 billion
soes_losses_filtered <- subset(soes_losses, Losses_Rs_Billion > 20)

# Create the treemap using treemapify
ggplot(soes_losses_filtered, aes(area = Losses_Rs_Billion, fill = SOE, label = paste(SOE, "\n", Losses_Rs_Billion, "B"))) +
  geom_treemap() +
  geom_treemap_text(fontface = "italic", colour = "black", place = "centre", grow = TRUE) +
  labs(title = "Treemap of SOEs Losses Greater than 20 Billion") +
  theme(legend.position = "none")  # Remove legend for simplicity

```


```{r}

# Create a data frame with SOEs and their corresponding losses
soes_losses <- data.frame(
  SOE = c("National Highway Authority (NHA)", 
          "Power Sector", 
          "Quetta Electric Supply Company Limited (QESCO)", 
          "Peshawar Electric Supply Company Limited (PESCO)", 
          "Lahore Electric Supply Company Limited (LESCO)", 
          "Sukkur Electric Power Company Limited (SEPCO)", 
          "Multan Electric Power Company Limited (MEPCO)", 
          "Hyderabad Electric Supply Company Limited (HESCO)", 
          "Faisalabad Electric Supply Company Limited (FESCO)", 
          "Islamabad Electric Supply Company Limited (IESCO)", 
          "Pakistan International Airlines Corporation (PIA)", 
          "Pakistan Railways", 
          "Pakistan Steel Mills Corporation (Private) Limited", 
          "Sui Southern Gas Company Limited (SSGC)", 
          "Pakistan Telecommunication Company Limited (PTCL)", 
          "GENCO-II Central Power Generation Company Limited, Thermal Power Station Guddu", 
          "GENCO-III Northern Power Generation Company Limited, Thermal Power Station Muzaffargarh", 
          "Pakistan Post Office", 
          "Pakistan Broadcasting Corporation (PBC)", 
          "GENCO-I Jamshoro Power Company Limited", 
          "GENCO-IV Lakhra Power Generation Company Limited", 
          "Pakistan Expo Centres (Pvt) Ltd", 
          "Pakistan Revenue Automation (Private) Limited", 
          "State Engineering Corporation (Private) Limited"),
  
  Losses_Rs_Billion = c(413.454, 
                        285, 
                        88.487, 
                        80.595, 
                        30.839, 
                        30.667, 
                        23.372, 
                        15.640, 
                        14.983, 
                        0.666, 
                        75.755, 
                        48.534, 
                        25.455, 
                        17.350, 
                        15.544, 
                        14.170, 
                        3.240, 
                        3.234, 
                        1.617, 
                        0.910, 
                        0.480, 
                        0.181, 
                        0.165, 
                        0.032)
)

# Arrange in descending order of losses and select only SOE and Losses_Rs_Billion columns
soes_losses <- soes_losses[order(-soes_losses$Losses_Rs_Billion), c("SOE", "Losses_Rs_Billion")]


# Optionally, for visual display in your R environment, use View() or save to CSV
# View(soes_losses)
# write.csv(soes_losses, "SOEs_Losses_Descending_Order.csv")
# Arrange in descending order of losses and select only SOE and Losses_Rs_Billion columns
soes_losses <- soes_losses[order(-soes_losses$Losses_Rs_Billion), c("SOE", "Losses_Rs_Billion")]

# Use knitr::kable to render a readable table with both columns visible side by side
knitr::kable(soes_losses, col.names = c("SOE", "Losses (Rs Billion)"), align = 'l')
```



## Introduction

- SOEs in Pakistan are underperforming. Why? Governance issues, lack of accountability, political interference, and corruption. 

- Debate on SOEs: privatization vs government shareholding. 

- Financial crisis demands a  reassessment of SOEs.  

- **Privatisation** is considered a key tool for **restructuring** and **improving efficiency** of SOEs. (Assumption)


## Arguments for Privatisation

- Privatisation can reduce the **fiscal burden** on the government.

- Privatisation can **improve efficiency** and **reduce losses** in SOEs. 

- Privatisation can **increase competition** and **improve service quality**. 

- Privatisation can **attract foreign investment** and **improve economic growth**. 

- Privatisation can **reduce corruption** and **improve transparency**. 

- Reduce burden on taxpayers 

## Arguments Against Privatisation

- Privatisation can lead to **job losses** and **wage cuts**. 

- Privatisation can lead to **monopolies** and **higher prices**. 
- Privatisation can lead to **loss of control** over strategic sectors. 

- Privatisation can lead to **foreign ownership** and **capital flight**. 

- Privatisation can lead to **inequality** and **social unrest**.    

## Key Principles of Privatisation 


- Efficiency & Profitability: The divestment should improve the operational efficiency of the enterprises.  

- **Transparency and fair pricing**: The process should be transparent and open to public scrutiny. 

- **Competition**: The market should be competitive and open to new entrants. 

- **Regulation**: The sector should be regulated to prevent abuse of market power. 

- **Social Impact**: The social impact of privatisation should be carefully considered. 

- **Economic Impact**: The economic impact of privatisation should be carefully considered. 

- **Political Impact**: The political impact of privatisation should be carefully considered. 

- **Legal Framework**: The legal framework for privatisation should be clear and well-defined. 

- **Public Support**: There should be public support for privatisation. 

## No one size fits all approach 

- Privatisation should be tailored to the specific circumstances of each SOE. 

- Different SOEs may require different approaches to privatisation. 

- The goal of privatisation should be to improve the performance of SOEs, not just to raise revenue or reduce the fiscal deficit. 

- Privatisation should be part of a broader strategy to reform the public sector and improve governance. 

- Privatisation should be accompanied by measures to protect workers and consumers. 


## Privatisation of Transport Sector

- Airports, airlines,  roads and railways are key components of the transport sector.

- **PIA** : Rs. 70 to 80 billion losses annually on accounting books. Economic loses are much higher.


## Privatisation of Power Sector

- Power generation, transmission and distribution are key components of the power sector. 

- **Circular debt** is a major issue in the power sector. 

- DISCOs are **inefficient** and **loss-making**.

- Each DISCO has its own set of problems and challenges.


## Success stories of privatisation 

- **Karachi Electric**: Privatisation of K-Electric has improved efficiency and service quality. 

- **PTCL**: Privatisation of PTCL has attracted foreign investment and improved service quality. 

- **HBL**: Privatisation of HBL has attracted foreign investment and improved service quality. 

- **UBL**: Privatisation of UBL has attracted foreign investment and improved service quality. 

Youth got chance to be anchors, ceos and managers in open tv and telcos.

## PTCL Privatisation

- Total employees: 61,000
- 2006: Etisalat buys 26% shares for $2.6 billion. 

- 2006: VSS offered to 30,000 employees.

- Second VSS offered in 2010 and almost 16000 employees left. 

- Etisallat invested $1.4 billion in PTCL. It has not paid $800 million yet just because of land title issue for assets which are with it but not in its name. 

## SOEs

80 SOEs are privatized in Pakistan. But there is no empirical evidence that in which sector privatisation is successful and in which sector it is not.

- No empirical evidence that privatisation is successful in all sectors. 

- Data are not available for all sectors.


## Example: Pakistan International Airlines & Pakistan Steel Mills
- These SOEs are incurring huge losses.
- Local and international competitors are more efficient.
- Divestment or privatization should be prioritized.

## Case of Pakistan Railways
- Vital for both **passenger traffic and freight services**.
- **Underinvestment** in Pakistan Railways.
- Need for a model similar to the **aviation sector**:
  - Government manages non-moving infrastructure.
  - Private sector manages moving infrastructure.

## Petroleum and Power Sectors
- Monopoly status of SOEs in these sectors.
- Profitability without benchmarking.
- **Rs 3 trillion circular debt** in the power sector due to inefficient governance.
- Need for market reforms before privatization in power distribution.

## Lessons from Karachi Electric
- Privatization alone does not solve all issues.
- **Sectoral reforms** needed for private sector efficiency gains.

## Need for Transparency
- Listing SOEs on the **stock exchange** can improve governance and transparency.
- Federal government’s **P3A Act** for public-private partnerships.

## Revisiting the Privatization Commission
- Legal structure for privatization needs to be updated.
- Need for **innovative divestment options** beyond traditional asset sales.

## Conclusion: Broadening the Debate
- Privatization debate should include:
  - What and how to privatize.
  - Efficient and transparent governance of SOEs.
  - Specialized treatment for each SOE.

[The Chaotic Pursuit of Privatisation of SOEs](https://www.dawn.com/news/1824841)

## Introduction
- Privatisation progress reported for **Pakistan International Airlines (PIA)** and outsourcing of three airports.
- **Privatisation Commission** working on a three-phase strategy to privatise SOEs.
- Focus on **loss-making SOEs** and reducing public sector involvement.

## Financial Burden of SOEs
> - SOEs incur annual losses of **Rs500 billion**.
> - **Total accumulated losses** exceed Rs2.5 trillion (approx. **$9 billion**).
- These losses pose a **systemic risk** to the financial sector.



## IMF and Privatisation
- **IMF bailout conditions** emphasize privatisation of PIA, Pakistan Steel Mills, RLNG power plants, and electricity distribution companies.
- Privatisation is part of the **structural reforms** under the IMF’s rescue loan package.

## Investment Expectations
- Special Investment Facilitation Council (SIFC) expects over **$50 billion** investment from UAE and Saudi Arabia in five years.
- So far, only a **fraction** of this investment has materialized in projects like Karachi Port.

## Historic Context of Privatisation
- Pakistan began privatising SOEs under the **IMF’s Structural Adjustment Programme** in the 1980s.
- Over the last three decades, privatisation has returned **Rs650 billion ($2.36 billion)** to the government.

## Success Stories of Privatisation
- **Banks, telecom industry, and electronic media** are often highlighted as successful examples of privatisation.
- Privatisation has saved recurring losses and improved efficiency in certain sectors.

## Cautions Against Hasty Privatisation
- Experts like **Haroon Sharif** caution against rushing into privatisation without a clear rationale.
- **World Bank** raises concerns about economic volatility, judicial activism, and trade union resistance as factors hindering successful privatisation.

## Need for Regulatory Reforms
- **Regulators need to be independent** and staffed with professionals to ensure competition and transparency.
- **Transaction structure** should promote long-term economic prosperity, avoiding monopolies.

## Risks of Monopolies
- Privatization may shift **monopolies from public to private sector**.
- Strong regulations are essential to prevent private sector monopolies from emerging after privatisation.

## Conclusion
- Privatisation requires extensive **reforms and transparency** to be successful.
- Without these, the progress will remain chaotic and may cause more harm than good.





# Real Estate

| S# | Name of PSEs |
|----|-------------------------------|
| 1  | Services International Hotel, Lahore |
| 2  | Jinnah Convention Centre, Islamabad |
| 3  | PIA-IL (Roosevelt Hotel NY) |
| 4  | Sale of Federal Government properties |

# Aviation Sector

| S# | Name of PSEs |
|----|-------------------------------------|
| 5  | Pakistan International Airlines Co. Ltd. |

# Financial Sector

| S# | Name of PSEs |
|----|-------------------------------------|
| 6  | Pakistan Re-Insurance Co. Ltd. (PakRe) |
| 7  | State Life Insurance Co. Ltd. |
| 8  | House Building Finance Corporation (HBFC) |
| 9  | First Women Bank Limited |

# Industrial Sector

| S# | Name of PSEs |
|----|-------------------------------------|
| 10 | Sindh Engineering Limited (SEL) |
| 11 | Pakistan Engineering Company (PECO) |

# Energy Sector

| S# | Name of PSEs |
|----|-------------------------------------|
| 12 | 1223 MW Balloki Power Plant |
| 13 | 1230 MW Haveli Bahadur Power Plant |
| 14 | Guddu Power Plant (747 MW) |
| 15 | Nandipur Power Plant (425 MW) |
| 16 | Faisalabad Electric Supply Company (FESCO) |
| 17 | Islamabad Electric Supply Company (IESCO) |
| 18 | Lahore Electric Supply Company (LESCO) |
| 19 | Gujranwala Electric Power Company (GEPCO) |
| 20 | Multan Electric Power Company (MEPCO) |
| 21 | Peshawar Electric Supply Company (PESCO) |
| 22 | Hyderabad Electric Supply Company (HESCO) |
| 23 | Quetta Electric Supply Company (QESCO) |
| 24 | Sukkur Electric Power Company (SEPCO) |
| 25 | Tribal Electric Supply Company (TESCO) |

# Issue with DISCOS privatisation

[DISCOS](https://profit.pakistantoday.com.pk/2024/06/19/discos-is-there-a-way-out-of-the-existential-quagmire/) 


## Overview
- Distribution companies (DISCOs) face persistent financial losses.
- Current efforts focus on selling DISCOS to private investors.
- The central question: Will privatization alone solve the core issues?

## Key Challenges
- **Electricity theft** and **low recovery rates** are major challenges.
- Losses concentrated in remote areas and urban slums.
- Privatization may not directly address the root causes.

## Role of Privatization
- K Electric as a case study shows mixed results.
- A need to determine if Transmission & Distribution (T&D) can sustain private investment alone.

## Proposed Solutions
- Distributed generation (solar) for remote areas.
- Increased electricity sales to reduce underutilized capacity.
- Governance and accountability should be devolved to local levels.

## Conclusion
- Strong **policy reforms** and **strategic privatization** are necessary.
- Privatization must focus on **separating high loss/low recovery areas** and creating a **competitive electricity market**.







# [SOE Privatisation](https://www.brecorder.com/news/40253974/soes-privatisation-the-only-solution)

## Introduction
Driven by the current ‘four-point must list of IMF,’ Pakistan’s government is focusing on the settlement of loss-making state-owned enterprises (SOEs). These SOEs have long been a concern for the IMF and a drain on Pakistan’s meager resources.

## Current State of SOEs
- **Circular debt**: Rs 2.7 trillion, despite repeated increases in electricity tariffs.
- **Loans and guarantees**: Nearly Rs 6 trillion to sustain loss-making SOEs.

Prime Minister Shehbaz Sharif has prioritized the reform of these institutions. There is a need for a modern system and expert leadership on the boards of these institutions to enhance their performance.

## New SOE Policy
The Ministry of Finance (MoF) has prepared a draft SOE policy to manage the fiscal risks. Key highlights of the policy:
- **SOEs in Pakistan**: Around 200, engaged in various activities.
- **Economic significance**: SOEs manage a significant share of Pakistan's infrastructure, communication, and utilities sectors.
- **Factors for optimal performance**:
  - Consistent overarching law for SOEs.
  - Clear SOE ownership and reform policy.
  - Commercial mandate and capable governance.
  - Centralized SOE ownership monitoring.
  - Strategic analysis and financial monitoring.
  - Accountability and transparency.

## Strategic SOEs
The draft policy identifies **strategic SOEs** that cannot be privatized due to their:
- Strategic, security, or social importance.
- Monopoly status without effective regulatory oversight.

## Challenges and Reforms
Successive governments have attempted SOE reforms, but with limited success. The key issues include:
- **SOEs beyond restructuring**: SOEs like PIA and Pakistan Steel Mills need massive investments in technology and human resources.
- **Political consequences**: Reforms often stall due to political considerations, despite the financial burden.




## International Example: India
India privatized its SOEs in the late 1990s. Key successes include:
- **Defense sector**: Majority of defense production is now in the private sector.
- **National carrier**: Air India was privatized and is being restructured by the Tata Group to compete globally.

# Demands from Government

## Key Demands

- **An independent board**: The board should be independent of political influence.

- **Forensic audit**: A forensic audit of SOEs should be conducted to identify the root causes of their financial distress in 3 months.

- **Reform plan**: A reform plan should be prepared for each SOE within 6 months.

## Advancements in Governance 

- **Technology, globalization, and management** have evolved.
- Traditional governance models of SOEs are outdated. 


## Tailored Approach Needed 

- **No one-size-fits-all** approach for privatization.
- Specialized treatment required for each SOE based on local and global circumstances.

## Privatisation of elite clubs

- **Privatization of elite clubs** like the Islamabad Club, Lahore Gymkhana, and Karachi Gymkhana can generate significant revenue. 

- **GORs**: Government Officers Residences (GORs) in prime locations can be sold for commercial purposes. 

- **Golf Courses**: Golf courses in prime locations can be privatized. 

- **State Guest Houses**: Punjab House, Sindh House ... land can be sold or leased for commercial purposes to hotels, shopping malls, etc.


# Introduction

Privatization is a widely debated concept. But what does it actually mean?

# Public Infrastructure and Investments

- Generations ago, America built:
  - National highways
  - Public colleges and universities
  - Public schools, parks, libraries, research, etc.
- Mid 1980, privatization was pushed to boost efficiency and reduce taxes.
  
> **Reality**: Privatization often boosts corporate profits instead.


> **Question**: Are these services reaching where they are most needed?

# The Role of Private Corporations

- Private corporations can be efficient.
- Some privatization has worked.
  
**BUT**: Their goal is to maximize **profit** for shareholders, not serve the **public interest**.

# The Public Interest

The key question: **What’s best for the public?**

Let’s consider 5 rules of thumb for when public services should not be privatized.

---

# 1. Community Services

- Don’t privatize when the service aims to bring communities together.
- Example: The **Railways and Roads** serves even the most rural areas.
- Privatization risks isolating marginalized groups.



# 5. Lack of Competition

- Don’t privatize where there’s insufficient competition.
- Example: **Defense contractors** face limited competition, leading to cost overruns.

> Privatization without competition results in inefficiency.



# The Public vs. Profit


